---
title: "Tips_dataset"
format: html
editor: visual
---

> Imagine you are the waiter at a restaurant and want to estimate the tip based on the meal and some other characteristics. The goal of this project is to build linear models to predict `tips` based on characteristics of the meals. Use Root Mean Squared Error (RMSE) as the metric of error. We are looking for the model with lowest RMSE on the `tesing` data.

```{r}
# RMSE
u <- c(1, 3, 2, -1, 5)
v <- c(1, 2, 4, -1.5, 3)
sqrt(mean((u-v)^2))         # RMSE
```


## Loading some important libraries

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)
```


## Importing dataset

We will use the dataset `tips` and for more information, read: [tips documentation](https://rdrr.io/cran/reshape2/man/tips.html).

```{r}
tip <- read_delim('https://raw.githubusercontent.com/mwaskom/seaborn-data/master/tips.csv', 
                  show_col_types = FALSE)
```

In order to work on the project, you should follow the following steps:

1. Study the distribution of each variable. Trying to understand them by computing some statistics: mean, variance, ... and plot some graphics.

2. Study the relation of each predictors to the target `tips` (statistically and graphically). Base on your investigation up to this stage, can you guess what variables are important to predict the tips?

3. Build linear models based on your choice of inputs using `lm` function. The model should be trained on the training data and predict the testing data as defined below.

```{r}
n <- nrow(tip)
train <- logical(n)
set.seed(168) # This will fix the train-test split. You will have the same train and test data
train[sample(n, 0.8*n)] <- TRUE
tip_train <- tip[train,]         # Train data
tip_test <- tip[!train,]         # Test data
```
4. Beautiful graphics and comments on your analysis are appreciated and to be expected. Good luck!
