---
title: "Chapter3: Linear regression"
author: "Sothea"
format: html
editor: visual
---

## Simple linear regression

```{r}
# install.packages('ISLR2')
# install.packages("tidyverse")
library(ISLR2)
library(tidyverse)
Wage %>%
  glimpse()


Wage %>%
  summary()
```



```{r}
y <- Wage$wage


Wage %>%
  select(is.numeric) %>%
  cor()



Wage %>%
  ggplot(aes(x = age,
             y = wage)) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y ~ x)



mod <- lm(wage~age, data = Wage)

summary(mod)
```


```{r}

df <- Wage %>%
  mutate(pred = mod$fitted.values)         # create new column on top of the dataset


df %>%
  ggplot(aes(x = age,
             y = wage)) +
  geom_point() +
  geom_smooth(method = "lm",
              formula = y ~ x) -> p

df_pred <- tibble(age = Wage$age,
                  wage = mod$fitted.values)

p +
  geom_point(data = df_pred, 
             color = "red",
             size = 2)
```


```{r}
x <- seq(-3,3, length.out = 100)
y <- dt(x, df = nrow(Wage) - 2)

tibble(x = x,
       density = y) %>%
  ggplot(aes(x = x, y  = density)) +
  geom_line(color = "blue") +
  geom_line(data = tibble(x = x,
                          density = dnorm(x)), color = "red")
```

## Homework

1. Count how many distinct values of variable `year`?

2. Make a graphic of relation between `year` and `wage`.

3. Is it more reasonable to look at year as a categorical variable? Transform `year` to be *factor* using `as_factor` function.

4. `lm` can predict `wage` using categorical variables. Build simple linear regression using other categorical variables (try each at a time against `wage`). What is the most interesting categorical variable in prediction `wage`?



